---
title: "Time Series Features: feasts"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: false
      number_sections: true
      fig_caption: TRUE
    css: "style.css"
editor_options: 
  chunk_output_type: console
---

<style>
.table-hover > tbody > tr:hover { 
  background-color: #f4f442;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(tsibble) #yearweek
#univariate and multivariate time series forecasting models, tsibble
library(fable) 

library(tidyverse)
library(here)
library(data.table)
library(lubridate)
library(dplyr)
library(patchwork)

#new library
library(GGally)
library(kableExtra)
library(here)

#getting all the trasformed data
source(here("source","get_fpp3.R"))
```

Reading: CH4.1 to 4.5

|Type | why?|
|:----|:-----|
|Short-term forecasts| needed for the scheduling of personnel, production and transportation. As part of the scheduling process, forecasts of demand are often also required.
|Medium-term forecasts |are needed to determine future resource requirements, in order to purchase raw materials, hire personnel, or buy machinery and equipment.|
|Long-term forecasts|used in strategic planning. Such decisions must take account of market opportunities, environmental factors and internal resources.|

- When we decompose a time series into components, we usually combine the trend and cycle into a single trend-cycle component (often just called the trend for simplicity). 

- Thus we can think of a time series as comprising three components: 
  - a trend-cycle component, 
  - a seasonal component, and 
  - a remainder component (containing anything else in the time series). 

- For some time series (e.g., those that are observed at least daily), there can be more than one seasonal component, corresponding to the different seasonal periods.

# Feast

- The feasts package includes functions for computing FEatures And Statistics from Time Series

```{r}
tourism |>
  features(Trips, list(mean = mean)) |> head() %>% 
  arrange(mean) %>% kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

tourism |> features(Trips, quantile) %>% head() %>%  
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## acf

|Column | Name |
|:------|:-----|
|acf1 |the first autocorrelation coefficient from the original data|
|acf10|the sum of squares of the first ten autocorrelation coefficients from the original data|
|diff1_acf1| the first autocorrelation coefficient from the differenced data|
|diff1_acf10| the sum of squares of the first ten autocorrelation coefficients from the differenced data|
|diff2_acf1|the first autocorrelation coefficient from the twice differenced data|
|diff2_acf10|the sum of squares of the first ten autocorrelation coefficients from the twice differenced data|
|season_acf1|For seasonal data, the autocorrelation coefficient at the first seasonal lag is also returned|

```{r}
tourism |> features(Trips, feat_acf) %>% head() %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## STL

|Statistic | Description |
|:-----|:------|
|$F_T$ | Strength of the trend between 0 and 1. Because the variance of the remainder might occasionally be even larger than the variance of the seasonally adjusted data, we set the minimal possible value of $F_T$ equal to zero|
|A series with seasonal strength $F_S$ close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1|

These measures can be useful, for example, when you have a large collection of time series, and you need to find the series with the most trend or the most seasonality.

```{r}
tourism |>
  features(Trips, feat_stl) %>% head() %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r}
tourism |>
  features(Trips, feat_stl) |>
  ggplot(aes(x = trend_strength, y = seasonal_strength_year,
             col = Purpose)) +
  geom_point() + theme(legend.position = "top") +
  facet_wrap(vars(State))
```

- Select region with maximum seasonal_stregth 

```{r}
tourism %>% head() %>%  
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

length(unique(tourism$Region))
length(unique(tourism$State))
length(unique(tourism$Purpose))
```


```{r}
temp <- tourism |>
  features(Trips, feat_stl) 

head(temp) %>%  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

#select the max seasonal_stregnth_year
temp <- temp %>% filter(
    seasonal_strength_year == max(seasonal_strength_year)) 

head(temp) %>%  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

- Join the tables

```{r}
temp <- temp %>% left_join(tourism, 
                   by = c("State", "Region", "Purpose"), 
                   multiple = "all")

head(temp)

temp %>%   ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

## Other statistics 

```{r}
#read in  excel file and display it as table
```

# Exploring Australian toursim data

```{r}
#it takes a few minutes
tourism_features <- tourism |>
  features(Trips, feature_set(pkgs = "feasts"))

glimpse(tourism_features)
```

```{r}
library(glue)
tourism_features |>
  select_at(vars(contains("season"), Purpose)) |>
  mutate(
    seasonal_peak_year = seasonal_peak_year +
      4*(seasonal_peak_year==0),
    seasonal_trough_year = seasonal_trough_year +
      4*(seasonal_trough_year==0),
    seasonal_peak_year = glue("Q{seasonal_peak_year}"),
    seasonal_trough_year = glue("Q{seasonal_trough_year}"),
  ) |>
  GGally::ggpairs(mapping = aes(colour = Purpose))
```

- It is difficult to explore more than a handful of variables in this way. 
- A useful way to handle many more variables is to use a dimension reduction technique such as principal components. 
  - This gives linear combinations of variables that explain the most variation in the original data. 
  - We can compute the principal components of the tourism features as follows.
  
```{r}
library(broom)
pcs <- tourism_features |>
  select(-State, -Region, -Purpose) |>
  prcomp(scale = TRUE) |>
  augment(tourism_features)

p1 <- pcs |>
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)


library(plotly)
ggplotly(p1)
```
  
- Each point represents one series and its location on the plot is based on all 48 features. 

- The first principal component (`.fittedPC1`) is the linear combination of the features which explains the most variation in the data. 

- The second principal component (`.fittedPC2`) is the linear combination which explains the next most variation in the data, while being uncorrelated with the first principal component. 

- The figure reveals a few things about the tourism data. 
  - First, the `holiday series` behave quite differently from the rest of the series. 
    - Almost all of the holiday series appear in the top half of the plot, while almost all of the remaining series appear in the bottom half of the plot.
    - the second principal component is distinguishing between holidays and other types of travel.

- The plot also allows us to identify anomalous time series — series which have unusual feature combinations. 

- These appear as points that are separate from the majority of series. 

- There are four that stand out, and we can identify which series they correspond to as follows

```{r}
outliers <- pcs |>
  filter(.fittedPC1 > 10) |>
  select(Region, State, Purpose, .fittedPC1, .fittedPC2)

outliers %>% kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r}
outliers |>
  left_join(tourism, by = c("State", "Region", "Purpose"), multiple = "all") |>
  mutate(Series = glue("{State}", "{Region}", "{Purpose}", .sep = "\n\n")) |>
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(Series ~ ., scales = "free") +
  labs(title = "Outlying time series in PC space")
```

We can speculate why these series are identified as unusual.

- Holiday visits to the south coast of NSW is highly seasonal but has almost no trend, whereas most holiday destinations in Australia show some trend over time.

- Melbourne is an unusual holiday destination because it has almost no seasonality, whereas most holiday destinations in Australia have highly seasonal tourism.

- The north western corner of Western Australia is unusual because it shows an increase in business tourism in the last few years of data, but little or no seasonality.

- The south western corner of Western Australia is unusual because it shows both an increase in holiday tourism in the last few years of data and a high level of seasonality.